---
title: "Recommendation system for Books"
author: "Reagan Kesseku"
output: html_document
---

```{r setup, include = FALSE}
# set global chunk options
# -------------------------
knitr::opts_chunk$set(echo = TRUE, cache = F, tidy = T,
                      warning = F, message = F, options(digits = 3))
```

```{r}
# set working directory
# ---------------------
setwd("D:/Ph.D_materials/Programming/R_programming/mdsr/Books-recommendation-systems")
```


```{r}
# Load functions and packages
# ---------------------------
source("book_pkg.R")
```

```{r}
# parallel computing
# ----------------------
cl = makePSOCKcluster(5)
registerDoParallel(cl)
```

#### Creating Books Recommendation System using Machine Learning algorithm in R
```{r}
# Import the datasets  
# -------------------
books <- read.csv("BX-Books.csv", header =  T, sep = ";")
users <- read.csv("BX-users.csv", header =  T, sep = ";")
ratings <- read.csv("BX-Book-Ratings.csv", header =  T, sep = ";")
```


```{r}
# Take a sample look at the book dataset
# --------------------------------------
books %>% sample_n(7) %>% datatable(options = list(scrollX = TRUE))
```

For the books dataset, there are `r dim(books)[1]` observations and `r dim(books)[2]` variables in the movies data. While the top five (5) column names (ISBN, Book title, Author, Year of publication, publisher) are clear book text or numeric data, the remaining 3 are images.


#### Anyway, in order to generate more realistic data, we will include a new variable called ‘Category’. This variable will indicate if the book belongs to any of the following categories:

Action and Adventure.
Classic.
Detective and Mystery.
Fantasy.

```{r}
set.seed(123)
categories = c("Action and Adventure", "Classic",
               "Detective and Mystery", "Fantasy")

books <- books %>% 
  mutate(category = sample(categories, n(), replace = TRUE, 
                           prob = c(0.25, 0.3, 0.25, 0.20)) %>% 
                    as.factor())

rm(categories)
```

```{r}
books <- books %>%
  mutate(ISBN = paste0("ISBN.", ISBN))
users <- users %>%
  mutate(User.ID = paste0("User.", User.ID))
ratings <- ratings %>%
  mutate(ISBN = paste0("ISBN.", ISBN),
         User.ID = paste0("User.", User.ID))

books %>% sample_n(2) %>% datatable(options = list(scrollX = TRUE))
```

Exploring the rating data
On the other hand, we will see how the ratings of the books are distributed. 
This is very important for the collaborative recommendation systems that we will build with R.

library(ggplot2)

```{r}
ratings %>%
  group_by(Book.Rating) %>%
  summarize(cases = n()) %>%
  mutate(percentage = cases / sum(cases) * 100) %>%
  ggplot(aes(Book.Rating, cases, fill = factor(Book.Rating))) + 
  geom_col() +
  geom_text(aes(label = paste0(round(percentage, 1), "%")), 
            position = position_stack(vjust = 0.5), 
            color = "black", size = 2.5) +
  scale_fill_brewer(palette = "Set3") +
  theme_bw() + 
  labs(x= "Book ratings", y = "Counts", fill = "Book ratings",
       title = "Distribution of book ratings for users") +
  scale_x_continuous(breaks = 0:10)

```

It is apparent that there are several zeros. 
This is not surprising as most people read books without rating them.
Thus, we will filter the complete ratings cases.

```{r}
# filter non zero ratings
# ------------------------
ratings <- ratings %>% filter(Book.Rating != 0)

ratings %>% dim()   # new dimension
```
```{r}
# Resulting distribution after filtering
# --------------------------------------
ratings %>%
  group_by(Book.Rating) %>%
  summarize(cases = n()) %>%
  mutate(percentage = cases / sum(cases) * 100) %>%
  ggplot(aes(Book.Rating, cases, fill = factor(Book.Rating))) + 
  geom_col() +
  geom_text(aes(label = paste0(round(percentage, 1), "%")), 
            position = position_stack(vjust = 0.5), 
            color = "black", size = 2.5) +
  scale_fill_brewer(palette = "Set3") +
  theme_bw() + 
  labs(x= "Book ratings", y = "Counts", fill = "Book ratings",
       title = "Distribution of book ratings for users") +
  scale_x_continuous(breaks = 0:10)
```
The change in percentages as displayed is a result of the total sample changing
after filtering the complete cases.

Finally, let’s see how much each person scores:

```{r}
# Users score
users_rate <- ratings %>%
  group_by(User.ID) %>% count() 

# provide summary measures
favstats(~ n, data = users_rate)
```

As we can see, 75% of users have given 3 recommendations or less. We are going to remove these people to keep only more significant users and, thus, reduce computing needs:

```{r}
user_index <- users_rate %>% filter(n > 4) %>% pull(User.ID)

users <- users %>%
         filter(User.ID %in% user_index)
ratings <- ratings %>%
         filter(User.ID %in% user_index)
books <- books %>%
         filter(ISBN %in% ratings$ISBN)
```


#### Content based Recommendation system

```{r}
# parallelized computing
# ----------------------
# cl = makePSOCKcluster(5)
# registerDoParallel(cl)

# Taking a subset of the data
book_feature <- books %>% 
  select(Book.Author, Publisher, category) %>% 
  slice(1:10000)

# convert to factors
book_feature <- book_feature %>% mutate(
  across(c(1, 2, 3), as.factor))

dissimilarity <- book_feature %>%
daisy(metric = "gower", weights = c(2, 0.5, 1)) %>%
as.matrix() 

row.names(dissimilarity)<-  books$ISBN[1:10000]
colnames(dissimilarity)<- books$ISBN[1:10000]

datatable(dissimilarity[15:20, 15:20])
```
Anyway, we are going to choose a user and keep the books they have read. We will apply the algorithm on these books:


```{r}
user_id = "User.1167"

user_books = ratings %>%
  filter(User.ID == user_id & ISBN %in% books$ISBN[1:10000]) %>%
  arrange(desc(Book.Rating))

head(user_books,10)
```

As we can see, the user has given 6 recommendations giving scores of 10, 9, 8, 7 and 5 points. Using the recommendations is something important, since it allows us to weigh the recommendations we give: we will prioritize a book similar to the one that has scored a 10 over one that has scored a 5. In addition, we will stop recommending those books that are similar to the which has scored a 0.


```{r}
books$ISBN = as.character(books$ISBN)
selected_books = user_books[ ,c("ISBN", "Book.Rating")]

recomendar = function(selected_books, dissimilarity_matrix,
                      books, n_recommendations = 5){

  selected_book_indexes = which(colnames(dissimilarity_matrix) %in% selected_books$ISBN)


  results = data.frame(dissimilarity_matrix[, selected_book_indexes],
                       recommended_book = row.names(dissimilarity_matrix),
                       stringsAsFactors = FALSE)


  recomendaciones = results %>%
    pivot_longer(cols = c(-"recommended_book") , names_to = "read_book",
                 values_to = "dissimilarity") %>%
      left_join(selected_books, by = c("recommended_book" = "ISBN"))%>%
    arrange(desc(dissimilarity)) %>%
    filter(recommended_book != read_book) %>%
    filter(!is.na(Book.Rating) ) %>%
    mutate(
      similarity = 1 - dissimilarity,
      weighted_score = similarity * Book.Rating) %>%
    arrange(desc(weighted_score)) %>%
    filter(weighted_score>0) %>%
    group_by(recommended_book) %>% slice(1) %>%
    top_n(n_recommendations, weighted_score)  %>%
    left_join(books, by = c("recommended_book" = "ISBN"))

  return(recomendaciones)
}

recomendaciones = recomendar(selected_books, dissimilarity, books)
datatable(recomendaciones, options = list(scrollX = TRUE))
```


```{r}
# # Create a directory to save the images
# dir <- "image_directory"
# dir.create(dir, showWarnings = FALSE)
# 
# # Data with URLs
# data <- read.table(text = "recommended_book	Image.URL.M
# ISBN.0060929596	http://images.amazon.com/images/P/0060929596.01.MZZZZZZZ.jpg
# ISBN.0373834985	http://images.amazon.com/images/P/0373834985.01.MZZZZZZZ.jpg
# ISBN.038533656X	http://images.amazon.com/images/P/038533656X.01.MZZZZZZZ.jpg
# ISBN.0440217520	http://images.amazon.com/images/P/0440217520.01.MZZZZZZZ.jpg
# ISBN.0679450548	http://images.amazon.com/images/P/0679450548.01.MZZZZZZZ.jpg
# ISBN.0743230213	http://images.amazon.com/images/P/0743230213.01.MZZZZZZZ.jpg
# ISBN.0802135196	http://images.amazon.com/images/P/0802135196.01.MZZZZZZZ.jpg",
#                    header = TRUE, 
# stringsAsFactors = FALSE)
# 
# 
# # For loop to download image
# for (i in 1:nrow(data)) {
#   url <- data[i, "Image.URL.M"]
#   filename <- paste0(dir, "/", basename(url))
#   download.file(url, filename, mode = "wb")
#   cat("Downloaded image", i, "out of", nrow(data), "\n")
# }
# 
# # Confirm the number of downloaded images
# num_images <- length(list.files(dir))
# cat("Downloaded", num_images, "images.\n")

```


```{r}
# visualizar_recomendacion = function(recomendation,
#                                      recommended_book, image, n_books = 5){
# 
#   if(n_books > nrow(recomendation)) {n_books = nrow(recomendation)}
# 
#   plot = list()
# 
#   if (!dir.exists("content_recommended_images")) {
#     dir.create("content_recommended_images")
#   }
# 
#   for(i in 1:n_books){
#     # Create dir & Download the images
#     img = pull(recomendation[i, colnames(recomendation) == image])
#     name = paste0("content_recommended_images/",i,".jpg")
#     suppressMessages(
#       download.file(as.character(img), destfile = name ,mode = "w") 
#     )
# 
#     download.file()
#     # Assign Objetc
#     plot[[i]] = rasterGrob(readJPEG(name))
#   }
# 
#   do.call(marrangeGrob, args = list(plot, ncol = n_books, nrow = 1, top=""))
# }
# 
# visualizar_recomendacion(recomendaciones, "recommended_book", "Image.URL.M")
```

